Simple repression draft

### 2020-09-26 doc/resubmission/section_04_bayesian_inference.tex
implement Rob's comments
implement Griffin's comments from 20200925

In this section of the paper, we continue our program of providing one complete
description of the entire broad sweep of studies that have been made in the
context of the repressor-operator model, dating all the way back to the original
work of Jacob and Monod and including the visionary quantitative work of
M\"{u}ller-Hill and collaborators~\cite{Oehler1990} and up to more recent
studies \cite{Garcia2011a}. In addition, the aim is to reconcile the equilibrium
and non-equilibrium perspectives that have been brought to bear on this problem.
From Section~\ref{section_02_means}, this reconciliation depends on a key
quantitative question as codified by Eq.~\ref{eq:deltaFR_eq_noneq_equiv}: does
the free energy of repressor binding, as described in the equilibrium models and
indirectly inferred from gene expression measurements, agree with the
corresponding values of repressor binding and unbinding rates in the
non-equilibrium picture, measured or inferred more directly? In this section we
tackle the statistical inference problem of inferring these repressor rates from
single-cell mRNA counts data. But before we can turn to the full case of simple
repression, we must choose an appropriate model of the constitutive promoter and
infer the parameter values in that model. This is the problem we address first.

@@

From consideration of Fano factors in the previous section, we suspect that
model 5 in Figure~\ref{fig2:constit_cartoons}(A), a one-state bursty model of
constitutive promoters, achieves the right balance of complexity and simplicity,
by allowing both Fano factor $\nu>1$, but also by remedying, by design, the
problems of parameter degeneracy that model 4 in
Figure~\ref{fig2:constit_cartoons} suffered~\cite{Razo-Mejia2020}. Does this
stand up to closer scrutiny, namely, comparison to full mRNA distributions
rather than simply their moments? We will test this thoroughly on single-cell
mRNA counts for different unregulated promoters from Jones et.\
al.~\cite{Jones2014}.

It will be instructive, however, to first consider the Poisson promoter, model 1
in Figure~\ref{fig2:constit_cartoons}. As we alluded to earlier, since the
Poisson distribution has a Fano factor $\nu$ strictly equal to 1, and all of the
observed data in Figure~\ref{fig2:constit_cartoons}(B) has Fano factor $\nu>1$,
we might already suspect that this model is incapable of fitting the data. We
will verify that this is in fact the case. Using the same argument we can
immediately rule out models 2 and 3 from Figure~\ref{fig2:constit_cartoons}(A).
These models have Fano factors $\nu\le 1$ meaning they are underdispersed
relative to the Poisson distribution. We will also not explicitly consider model
4 from~\fig{fig2:constit_cartoons}(A) since it was already thoroughly analyzed
in~\cite{Razo-Mejia2020}, and since model 5 can be viewed as a special case of
it.

Our objective for this section will then be to assess whether or not model 5 is
quantitatively able to reproduce experimental data. In other words, if our claim
is that the level of coarse graining in this model is capable of capturing the
relevant features of the data, then we should be able to find values for the
model parameters that can match theoretical predictions with single-molecule
mRNA count distributions. A natural language for this parameter inference
problem is that of Bayesian probability. We will then build a Bayesian inference
pipeline to fit the model parameters to data. To gain intuition on how this
analysis is done we will begin with the ``wrong'' model 1 in
Figure~\ref{fig2:constit_cartoons}(A). We will use the full dataset of
single-cell mRNA counts from~\cite{Jones2014} used in
Figure~\ref{fig2:constit_cartoons}(B).

@@

For this model the master equation of interest is
Eq.~\ref{eq:poisson_promoter_cme} with repressor set to zero, i.e.,
\begin{equation}
{d\over dt}p_U(m)(t) = 
        rp_U(m-1)(t) 
        - rp_U(m)(t)
        + (m+1)\gamma p_U(m+1)(t) 
        - \gamma p_U(m)(t),
\end{equation}
whose steady-state solution is given by a Poisson distribution with parameter
$\lambda \equiv r / \gamma$~\cite{Sanchez2013}. The goal of our inference 
problem is then to find the probability distribution for the parameter value
$\lambda$ given the experimental data. By Bayes' theorem this can be written as
\begin{equation}
p(\lambda \mid D) = {p(D \mid \lambda) p(\lambda) \over p(D)},
\end{equation}
where $D = \{m_1, m_2, \ldots, m_N \}$ are the single-cell mRNA experimental
counts. As is standard we will neglect the denominator $p(D)$ on the right
hand side since it is independent of $\lambda$ and serves only as a
normalization factor.

@@

\begin{figure}%[h!]
\centering
\includegraphics[width=\textwidth]{../figures/main/fig03.pdf}
\caption{\textbf{Constitutive promoter posterior inference and model
comparison.} (A) The joint posterior density of model 5, the bursty promoter
with negative binomially-distributed steady state, is plotted with MCMC samples.
1D marginal probability densities are plotted as flanking histograms. The model
was fit on \textit{lacUV5} data from~\cite{Jones2014}. (B) The empirical
distribution function (ECDF) of the observed population distribution of mRNA
transcripts under the control of a constitutive \textit{lacUV5} promoter is
shown in black. The median posterior predictive ECDFs for models (1), Poisson,
and (5), negative binomial, are plotted in dark green and dark blue,
respectively. Lighter green and blue regions enclose 95\% of all posterior
predictive samples from their respective models. Model (1) is in obvious
contradiction with the data while model (5) is not. Single-cell mRNA count data
is again from~\cite{Jones2014}. (C) Joint posterior distributions for burst rate
$k_i$ and mean burst size $b$ for 18 unregulated promoters
from~\cite{Jones2014}. Each contour indicates the 95\% highest posterior
probability density region for a particular promoter. Note that the vertical
axis is shared with (D). (D) Plots of the burst rate $k_i$ vs.\ the binding
energy for each promoter as predicted in~\cite{Brewster2012}. The dotted line
shows the predicted slope according to Eq.~\ref{eq:bursty_equil_corresp1},
described in text. Each individual promoter is labeled with a unique number in
both (C) and (D) for cross comparison and for comparison with 
Figure~\ref{fig2:constit_cartoons}(B).}
\label{fig:constit_post_full}
\end{figure}

@@

\subsubsection{Connecting inferred parameters to prior work}
It is interesting to connect these inferences on $k_i$ and $b$ to the work
of~\cite{Brewster2012}, where these same 18 promoters were considered through
the lens of the three-state equilibrium model (model 2 in
Figure~\ref{fig1:means_cartoons}(B)) and binding energies $\Delta\varepsilon_P$
were predicted from an energy matrix model derived from~\cite{Kinney2010}. As
previously discussed the thermodynamic models of gene regulation can only make
statements about the mean gene expression. This implies that we can draw the
connection between both frameworks by equating the mean mRNA $\left\langle m
\right\rangle$. This results in
\begin{equation}
\langle m \rangle = \frac{k_i b}{\gamma}
        = \frac{r}{\gamma}
        \frac{\frac{P}{N_{NS}}\exp(-\beta\Delta\varepsilon_P)}
                {1+\frac{P}{N_{NS}}\exp(-\beta\Delta\varepsilon_P)}.
\end{equation}
By taking the weak promoter approximation for the equilibrium model ($P/N_{NS} 
\exp(-\beta\Delta\varepsilon_r) \ll 1$) results in
\begin{equation}
\langle m \rangle = \frac{k_i b}{\gamma}
        = \frac{r}{\gamma} \frac{P}{N_{NS}}\exp(-\beta\Delta\varepsilon_P),
\end{equation}
valid for all the binding energies considered here.

@@

We mention one further comparison, between our inferred parameters and the
work of Chong et.\ al.~\cite{Chong2014}, which is interesting and puzzling.
Beautiful experiments in~\cite{Chong2014} convincingly argue that
supercoiling accumulated from the production of mRNA transcripts is key
in setting the burstiness of mRNA production.
In their model, this supercoiling occurs on the scale of $\sim100$~kb
domains of DNA. This suggests that all genes on a given domain should burst
in synchrony, and that the difference between highly and lowly expressed
genes is the \textit{size} of transcriptional bursts, not the
\textit{time between} bursts.
But here, all burst sizes we infer in Figure~\ref{fig:constit_post_full}(C)
are comparable and burst rates vary wildly. It is not immediately clear
how to square this circle. Furthermore, Figure~7E in~\cite{Chong2014}
reports values of the quantity they label $\beta/\alpha$ and we label
$k^+/k^-$ in model 4 from Figure~\ref{fig2:constit_cartoons}. In contrast
to the findings of~\cite{Razo-Mejia2020}, Chong et.\ al.\ do not find
$k^+/k^-\ll1$ for most of the genes they consider. This begs the question:
is the \textit{galK} chromosomal locus used for the reporter constructs
in~\cite{Razo-Mejia2020} and~\cite{Jones2014} merely an outlier,
or is there a deeper puzzle here waiting to be resolved? Without more
apples-to-apples data we can only speculate, and we leave it as an
intriguing open question for the field.

Despite such puzzles, our goal here is not to unravel the mysterious origins
of burstiness in transcription. Our remaining task in this work is a
determination of the physical reality of equilibrium binding energies in
Figure~\ref{fig1:means_cartoons}, as codified by the equilibrium-nonequilibrium
equivalence of Eq.~\ref{eq:deltaFR_eq_noneq_equiv}.
For our phenomenological needs here model 5 in
Figure~\ref{fig2:constit_cartoons} is more than adequate: the posterior
distributions in Figure~\ref{fig:constit_post_full}(C) are cleanly
identifiable and the predictive checks in Figure~\ref{figS:ppc_unreg}
indicate no discrepancies between the model and the mRNA single-moleucle
count data of~\cite{Jones2014}. Of the models we have considered it
is unique in satisfying both these requirements. So we will happily use
it as a foundation to build upon in the next section when we add regulation.

@@

Now that we have a satisfactory model in hand for constitutive promoters, we
would like to return to the main thread: can we reconcile the equilibrium and
nonequilibrium models by putting to the test
Eq.~\ref{eq:deltaFR_eq_noneq_equiv}, the correspondence between indirectly
inferred equilibrium binding energies and nonequilibrium kinetic rates? To make
this comparison, is it possible to infer repressor binding and unbinding rates
from mRNA distributions over a population of cells as measured by
single-molecule Fluorescence \textit{in situ} Hybridization in~\cite{Jones2014}?
If so, how do these inferred rates compare to direct single-molecule
measurements such as from~\cite{Hammar2014} and to binding energies such as
from~\cite{Garcia2011a} and~\cite{Razo-Mejia2018}, which were inferred under the
assumptions of the equilibrium models in Figure~\ref{fig1:means_cartoons}(B)?
And can this comparison shed light on the unreasonable effectiveness of the
equilibrium models, for instance, in their application in~\cite{Chure2019,
Chure2019a}?

\begin{figure}%[h!]
\centering
\includegraphics[width=\textwidth]{../figures/main/fig04.pdf}
\caption{\textbf{Simple repression parameter inference and comparison.}
(A) Contours which enclose 50\% and 95\% of the posterior probability mass are
shown for each of several 2D slices of the 9D posterior distribution. The model
assumes one unbinding rate for each operator (Oid, O1, O2) and one binding rate
for each aTc induction concentration (corresponding to an unknown mean repressor
copy number).
(B, upper) Ratios of our inferred unbinding rates are
compared with operator binding energy differences measured by Garcia and
Phillips~\cite{Garcia2011a} (triangles) and Razo-Mejia et.\
al.~\cite{Razo-Mejia2018} (circles). Blue glyphs compare O2-O1, while orange
compare O1-Oid. Points with perfect agreement would lie on the dotted line.
(B, lower) Unbinding rates for O1 (cyan) and Oid (red)
inferred in this work are compared with single-molecule measurements
from Hammar et.\ al.~\cite{Hammar2014}. We plot the comparison assuming illustrative mRNA
lifetimes of $\gamma^{-1}=3$~min (triangles) or $\gamma^{-1}=5$~min
(squares). Dotted line is as in upper panel.
(C) Theory-experiment comparison are shown for each of the datasets
used in the inference of the model in (A). Observed single-molecule mRNA counts
data from~\cite{Jones2014} are plotted as black lines. The median of the randomly
generated samples for each condition is plotted as a dark colored line. Lighter
colored bands enclose 95\% of all samples for a given operator/repressor copy
number pair.
%Samples are sorted by operator and plotted separately for visual clarity.
The unregulated promoter, \textit{lacUV5}, is shown with each as a reference.}
\label{fig4:repressed_post_full}
\end{figure}

@@

Before showing the mathematical formulation of our statistical inference model,
we would like to sketch the intuitive structure. The dataset
from~\cite{Jones2014} we consider consists of single-cell mRNA counts data of
nine different conditions, spanning several combinations of three unique
repressor binding sites and four unique repressor copy numbers. We assume that
the values of $k_i$ and $b$ are known, since we have already cleanly inferred
them from constitutive promoter data, and further we assume that these values
are the same across datasets with different repressor binding sites and copy
numbers. In other words, we assume that the regulation of the transcription
factor does not affect the mean burst size nor the burst initiation rate. The
regulation occurs as the promoter is taken away from the transcriptionally
active state when the promoter is bound by repressor. We assume that there is
one unbinding rate parameter for each repressor binding site, and likewise one
binding rate for each unique repressor copy number. This makes our model seven
dimensional, or nine if one counts $k_i$ and $b$ as well. Note that we use only
a subset of the datasets from Jones et.\ al.~\cite{Jones2014}, as discussed more
in Appendix~\ref{sec:bayesian}.

Formally now, denote the set of seven repressor rates to be inferred as
\begin{equation}
\vect{k} =\{k_{Oid}^-, k_{O1}^-, k_{O2}^-,
k_{0.5}^+, k_{1}^+, k_{2}^+, k_{10}^+\},
\end{equation}
where subscripts for dissociation rates $k^-$  indicate the different repressor
binding sites, and subscripts for association rates $k^+$ indicate the
concentration of the small-molecule that controlled the expression of the LacI
repressor (see Appendix~\ref{sec:bayesian}). This is because for this particular
dataset the repressor copy numbers were not measured directly, but it is safe to
assume that a given concentration of the inducer resulted in a specific mean
repressor copy number~\cite{Chure2019a}. Also note that the authors
of~\cite{Jones2014} report estimates of LacI copy number per cell rather than
direct measurements. However, these estimates were made assuming the validity of
the equilibrium models in Figure~\ref{fig1:means_cartoons}, and since testing
these models is our present goal, it would be circular logic if we were to make
the same assumption. Therefore we will make no assumptions about the LacI copy
number for a given inducer concentrations.

@@

\subsubsection{Comparison with prior measurements of repressor binding energies}
Our primary goal in this work is to reconcile the kinetic and equilibrium
pictures of simple repression. Towards this end we would like to compare the
repressor kinetic rates we have inferred with the repressor binding energies
inferred through multiple methods in~\cite{Garcia2011a}
and~\cite{Razo-Mejia2018}. If the agreement is close, then it suggests that the
equilibrium models are not wrong and the repressor binding energies they contain
correspond to physically real free energies, not mere fit parameters.

>>>>

In this section we aim to reconcile the thermodynamic and kinetic perspectives
that have been brought to bear on the repressor-operator model. From
Section~\ref{section_02_means}, this reconciliation depends on a key
quantitative question as codified by Eq.~\ref{eq:deltaFR_eq_noneq_equiv}: does
the free energy of repressor binding, as described in the thermodynamic models
and indirectly inferred from gene expression measurements, agree with the
corresponding values of repressor binding and unbinding rates in the kinetic
picture, measured or inferred more directly? In this section we tackle the
statistical inference problem of inferring these repressor rates from
single-cell mRNA counts data. But before we can turn to the full case of simple
repression, we must choose an appropriate model of the constitutive promoter and
infer the parameter values in that model. This is the problem we address first.

@@

From consideration of Fano factors in the previous section, we suspect that
model 5 in Figure~\ref{fig2:constit_cartoons}(A), a one-state bursty model of
constitutive promoters, achieves the right balance of complexity and simplicity,
by allowing both Fano factor $\nu>1$, but also by remedying, by design, the
problems of parameter degeneracy that model 4 in
Figure~\ref{fig2:constit_cartoons}(A) suffered~\cite{Razo-Mejia2020}. We will
test this thoroughly on single-cell mRNA counts for different unregulated
promoters from Jones et.\ al.~\cite{Jones2014}, but first we consider It will be
instructive, however, to first consider the Poisson promoter, model 1 in
Figure~\ref{fig2:constit_cartoons}. As we alluded to earlier, since the Poisson
distribution has a Fano factor $\nu$ strictly equal to 1, and all of the
observed data in Figure~\ref{fig2:constit_cartoons}(B) has Fano factor $\nu>1$,
we might already suspect that this model is incapable of fitting the data. We
will verify that this is in fact the case. Using the same argument we can
immediately rule out models 2 and 3 from Figure~\ref{fig2:constit_cartoons}(A).
These models have Fano factors $\nu\le 1$ meaning they are underdispersed
relative to the Poisson distribution. We will also not explicitly consider model
4 from~\fig{fig2:constit_cartoons}(A) since it was already thoroughly analyzed
in~\cite{Razo-Mejia2020}, and since model 5 can be viewed as a special case of
it.

\mrm{Griffin also feels this paragraph should go. Too many times reminding the 
reader what we are up to. I think he is right.}
Our objective for this section will then be to assess whether or not model 5 is
quantitatively able to reproduce experimental data. In other words, if our claim
is that the level of coarse graining in this model is capable of capturing the
relevant features of the data, then we should be able to find values for the
model parameters that can match theoretical predictions with single-molecule
mRNA count distributions. A natural language for this parameter inference
problem is that of Bayesian probability. We will then build a Bayesian inference
pipeline to fit the model parameters to data. To gain intuition on how this
analysis is done we will begin with the ``wrong'' model 1 in
Figure~\ref{fig2:constit_cartoons}(A). We will use the full dataset of
single-cell mRNA counts from~\cite{Jones2014} used in
Figure~\ref{fig2:constit_cartoons}(B).

@@

Model 1 in Figure~\ref{fig2:constit_cartoons}(A) predicts a  mRNA distribution
whose steady-state solution is given by a Poisson distribution with parameter
$\lambda \equiv r / \gamma$, where $r$ is the mRNA production rate, and $\gamma$
is the mRNA degradation rate~\cite{Sanchez2013}. The goal of our inference
problem is then to find the probability distribution for the parameter value
$\lambda$ given the experimental data. By Bayes' theorem this can be written as
\begin{equation}
p(\lambda \mid D) = {p(D \mid \lambda) p(\lambda) \over p(D)},
\end{equation}
where $D = \{m_1, m_2, \ldots, m_N \}$ are the single-cell mRNA experimental
counts. As is standard we will neglect the denominator $p(D)$ on the right
hand side since it is independent of $\lambda$ and serves only as a
normalization factor.

@@

\begin{figure}%[h!]
\centering
\includegraphics[width=\textwidth]{../../figures/main/fig03.pdf}
\caption{\textbf{Constitutive promoter posterior inference and model
comparison.} (A) The joint posterior density of model 5, the bursty promoter
with negative binomially-distributed steady state, is plotted with MCMC samples.
1D marginal probability densities are plotted as flanking histograms. The model
was fit on \textit{lacUV5} data from~\cite{Jones2014}. (B) The empirical
distribution function (ECDF) of the observed population distribution of mRNA
transcripts under the control of a constitutive \textit{lacUV5} promoter is
shown in black. The median posterior predictive ECDFs for models (1), Poisson,
and (5), negative binomial, are plotted in dark green and dark blue,
respectively. Lighter green and blue regions enclose 95\% of all posterior
predictive samples from their respective models. Model (1) is in obvious
contradiction with the data while model (5) is not. Single-cell mRNA count data
is again from~\cite{Jones2014}. (C) Joint posterior distributions for burst rate
$k_i$ and mean burst size $b$ for 18 unregulated promoters
from~\cite{Jones2014}. Each contour indicates the 95\% highest posterior
probability density region for a particular promoter. Note that the vertical
axis is shared with (D). (D) Plots of the burst rate $k_i$ vs.\ the binding
energy for each promoter as predicted in~\cite{Brewster2012}. The dotted line
shows the predicted slope according to Eq.~\ref{eq:bursty_equil_corresp1},
described in text. Each individual promoter is labeled with a unique number in
both (C) and (D) for cross comparison and for comparison with 
Figure~\ref{fig2:constit_cartoons}(B).}
\label{fig:constit_post_full}
\end{figure}

@@

\subsubsection{Connecting inferred parameters to prior work}
It is interesting to connect these inferences on $k_i$ and $b$ to the work
of~\cite{Brewster2012}, where these same 18 promoters were considered through
the lens of the three-state thermodynamic model (model 2 in
Figure~\ref{fig1:means_cartoons}(B)) and binding energies $\Delta\varepsilon_P$
were predicted from an energy matrix model derived from~\cite{Kinney2010}. As
previously discussed the thermodynamic models of gene regulation can only make
statements about the mean gene expression. This implies that we can draw the
connection between both frameworks by equating the mean mRNA $\left\langle m
\right\rangle$. This results in
\begin{equation}
\langle m \rangle = \frac{k_i b}{\gamma}
        = \frac{r}{\gamma}
        \frac{\frac{P}{N_{NS}}\exp(-\beta\Delta\varepsilon_P)}
                {1+\frac{P}{N_{NS}}\exp(-\beta\Delta\varepsilon_P)}.
\end{equation}
By taking the weak promoter approximation for the equilibrium model ($P/N_{NS} 
\exp(-\beta\Delta\varepsilon_r) \ll 1$) results in~\cite{Brewster2012}
\begin{equation}
\langle m \rangle = \frac{k_i b}{\gamma}
        = \frac{r}{\gamma} \frac{P}{N_{NS}}\exp(-\beta\Delta\varepsilon_P),
\end{equation}
valid for all the binding energies considered here.

@@

\mrm{ I would vote to remove this paragraph completely.}
We mention one further comparison, between our inferred parameters and the work
of Chong et.\ al.~\cite{Chong2014}, which is interesting and puzzling.
Experiments in~\cite{Chong2014} convincingly argue that supercoiling accumulated
from the production of mRNA transcripts is key in setting the burstiness of mRNA
production. In their model, this supercoiling occurs on the scale of
$\sim100$~kb domains of DNA. This suggests that all genes on a given domain
should burst in synchrony, and that the difference between highly and lowly
expressed genes is the \textit{size} of transcriptional bursts, not the
\textit{time between} bursts. But here, all burst sizes we infer in
Figure~\ref{fig:constit_post_full}(C) are comparable and burst rates vary
wildly. It is not immediately clear how to square this circle. Furthermore,
Figure~7E in~\cite{Chong2014} reports values of the quantity they label
$\beta/\alpha$ and we label $k^+/k^-$ in model 4 from
Figure~\ref{fig2:constit_cartoons}. In contrast to the findings
of~\cite{Razo-Mejia2020}, Chong et.\ al.\ do not find $k^+/k^-\ll1$ for most of
the genes they consider. This begs the question: is the \textit{galK}
chromosomal locus used for the reporter constructs in~\cite{Razo-Mejia2020}
and~\cite{Jones2014} merely an outlier, or is there a deeper puzzle here waiting
to be resolved? Without more apples-to-apples data we can only speculate, and we
leave it as an intriguing open question for the field.

Despite such puzzles, our goal here is not to unravel the mysterious origins of
burstiness in transcription. Our remaining task in this work is a determination
of the physical reality of thermodynamic binding energies in
Figure~\ref{fig1:means_cartoons}, as codified by the thermodynamic-kinetic
equivalence of Eq.~\ref{eq:deltaFR_eq_noneq_equiv}. For our phenomenological
needs here model 5 in Figure~\ref{fig2:constit_cartoons}(A) is more than
adequate: the posterior distributions in Figure~\ref{fig:constit_post_full}(C)
are cleanly identifiable and the predictive checks in
Figure~\ref{figS:ppc_unreg} indicate no discrepancies between the model and the
mRNA single-molecule count data of~\cite{Jones2014}. Of the models we have
considered it is unique in satisfying both these requirements. So we will
happily use it as a foundation to build upon in the next section when we add
regulation.

@@

Now that we have a satisfactory model in hand for constitutive promoters, we
would like to return to the main thread: can we reconcile the thermodynamic and
kinetic models by putting to the test Eq.~\ref{eq:deltaFR_eq_noneq_equiv}, the
correspondence between indirectly inferred thermodynamic binding energies and
kinetic rates? To make this comparison, is it possible to infer repressor
binding and unbinding rates from mRNA distributions over a population of cells
as measured by single-molecule Fluorescence \textit{in situ} Hybridization
in~\cite{Jones2014}? If so, how do these inferred rates compare to direct
single-molecule measurements such as from~\cite{Hammar2014} and to binding
energies such as from~\cite{Garcia2011a} and~\cite{Razo-Mejia2018}, which were
inferred under the assumptions of the thermodynamic models in
Figure~\ref{fig1:means_cartoons}(B)? And can this comparison shed light on the
unreasonable effectiveness of the equilibrium models, for instance, in their
application in~\cite{Chure2019, Chure2019a}?

\begin{figure}%[h!]
\centering
\includegraphics[width=\textwidth]{../../figures/main/fig04.pdf}
\caption{\textbf{Simple repression parameter inference and comparison.}
(A) Contours which enclose 50\% and 95\% of the posterior probability mass are
shown for each of several 2D slices of the 9D posterior distribution. The model
assumes one unbinding rate for each operator (Oid, O1, O2) and one binding rate
for each aTc induction concentration (corresponding to an unknown mean repressor
copy number).
(B, upper) Ratios of our inferred unbinding rates are
compared with operator binding energy differences measured by Garcia and
Phillips~\cite{Garcia2011a} (triangles) and Razo-Mejia et.\
al.~\cite{Razo-Mejia2018} (circles). Blue glyphs compare O2-O1, while orange
compare O1-Oid. Points with perfect agreement would lie on the dotted line.
(B, lower) Unbinding rates for O1 (cyan) and Oid (red)
inferred in this work are compared with single-molecule measurements
from Hammar et.\ al.~\cite{Hammar2014}. We plot the comparison assuming illustrative mRNA
lifetimes of $\gamma^{-1}=3$~min (triangles) or $\gamma^{-1}=5$~min
(squares). Dotted line is as in upper panel.
(C) Theory-experiment comparison are shown for each of the datasets
used in the inference of the model in (A). Observed single-molecule mRNA counts
data from~\cite{Jones2014} are plotted as black lines. The median of the randomly
generated samples for each condition is plotted as a dark colored line. Lighter
colored bands enclose 95\% of all samples for a given operator/repressor copy
number pair.
%Samples are sorted by operator and plotted separately for visual clarity.
The unregulated promoter, \textit{lacUV5}, is shown with each as a reference.}
\label{fig4:repressed_post_full}
\end{figure}

@@

Before showing the mathematical formulation of our statistical inference model,
we would like to sketch the intuitive structure. The dataset
from~\cite{Jones2014} we consider consists of single-cell mRNA counts data of
nine different conditions, spanning several combinations of three unique
repressor binding sites (the so-called $Oid$, $O1$, and $O2$ operators) and four
unique repressor copy numbers. We assume that the values of $k_i$ and $b$ are
known, since we have already cleanly inferred them from constitutive promoter
data, and further we assume that these values are the same across datasets with
different repressor binding sites and copy numbers. In other words, we assume
that the regulation of the transcription factor does not affect the mean burst
size nor the burst initiation rate. The regulation occurs as the promoter is
taken away from the transcriptionally active state when the promoter is bound by
repressor. We assume that there is one unbinding rate parameter for each
repressor binding site, and likewise one binding rate for each unique repressor
copy number. This makes our model seven dimensional, or nine if one counts $k_i$
and $b$ as well. Note that we use only a subset of the datasets from Jones et.\
al.~\cite{Jones2014}, as discussed more in Appendix~\ref{sec:bayesian}.

Formally now, denote the set of seven repressor rates to be inferred as
\begin{equation}
\vect{k} =\{k_{Oid}^-, k_{O1}^-, k_{O2}^-,
k_{0.5}^+, k_{1}^+, k_{2}^+, k_{10}^+\},
\end{equation}
where subscripts for dissociation rates $k^-$  indicate the different repressor
binding sites, and subscripts for association rates $k^+$ indicate the
concentration of the small-molecule that controlled the expression of the LacI
repressor (see Appendix~\ref{sec:bayesian}). This is because for this particular
dataset the repressor copy numbers were not measured directly, but it is safe to
assume that a given concentration of the inducer resulted in a specific mean
repressor copy number~\cite{Chure2019a}. Also note that the authors
of~\cite{Jones2014} report estimates of LacI copy number per cell rather than
direct measurements. However, these estimates were made assuming the validity of
the thermodynamic models in Figure~\ref{fig1:means_cartoons}, and since testing
these models is our present goal, it would be circular logic if we were to make
the same assumption. Therefore we will make no assumptions about the LacI copy
number for a given inducer concentration.

@@

\subsubsection{Comparison with prior measurements of repressor binding energies}
Our primary goal in this work is to reconcile the kinetic and thermodynamic
pictures of simple repression. Towards this end we would like to compare the
repressor kinetic rates we have inferred with the repressor binding energies
inferred through multiple methods in~\cite{Garcia2011a}
and~\cite{Razo-Mejia2018}. If the agreement is close, then it suggests that the
thermodynamic models are not wrong and the repressor binding energies they
contain correspond to physically real free energies, not mere fit parameters.

Figure~\ref{fig4:repressed_post_full}(B) shows both comparisons, with the top
panel comparing to thermodynamic binding energies and the bottom panel comparing
to single-molecule measurements. First consider the top panel and its comparison
between repressor kinetic rates and binding energies. As described in
section~\ref{section_02_means}, if the equilibrium binding energies
from~\cite{Garcia2011a} and~\cite{Razo-Mejia2018} indeed are the physically real
binding energies we believe them to be, then they should be related to the
repressor kinetic rates via Eq.~\ref{eq:deltaFR_eq_noneq_equiv}, which we
restate here,
\begin{equation}
\Delta F_R = \beta\Delta\varepsilon_R - \log(R/N_{NS})
        = - \log(k_R^+/k_R^-).
\label{eq:deltaFR_eq_noneq_equiv_repeat}
\end{equation}
Assuming mass action kinetics implies that $k_R^+$ is proportional to repressor
copy number $R$, or more precisely, it can be thought of as repressor copy
number times some intrinsic per molecule association rate. But since $R$ is not
directly known for our data from~\cite{Jones2014}, we cannot use this equation
directly. Instead we can consider two different repressor binding sites and
compute the \textit{difference} in binding energy between them, since this
difference depends only on the unbinding rates and not on the binding rates.
This can be seen by evaluating Eq.~\ref{eq:deltaFR_eq_noneq_equiv_repeat} for
two different repressor binding sites, labeled (1) and (2), but with the same
repressor copy number $R$, and taking the difference to find
\begin{equation}
\Delta F_R^{(1)} - \Delta F_R^{(2)}
= \beta\Delta\varepsilon_1 - \beta\Delta\varepsilon_2
= - \log(k_R^+/k_1^-) + \log(k_R^+/k_2^-),
\end{equation}
or simply
\begin{equation}
\beta\Delta\varepsilon_1 - \beta\Delta\varepsilon_2
= \log(k_2^-/k_1^-).
\end{equation}
The left and right hand sides of this equation are exactly the horizontal and
vertical axes of the top panel of Figure~\ref{fig4:repressed_post_full}. Since
we inferred rates for three repressor binding sites (O1, O2, and Oid), there are
only two independent differences that can be constructed, and we arbitrarily
chose to plot O2-O1 and O1-Oid in Figure~\ref{fig4:repressed_post_full}(B).
Numerically, we compute values of $k_{O1}^- / k_{Oid}^-$ and $k_{O2}^- /
k_{O1}^-$ directly from our full posterior samples, which conveniently provides
uncertainties as well, as detailed in Appendix~\ref{sec:bayesian}.
% \mmnote{Explain in appendix how marginalization with sampling makes this
% ``trivial'' to compute from samples of the joint posterior, b/c sampling makes
% marginalization trivial.}
We then compare these log ratios of rates to the
binding energy differences $\Delta\varepsilon_{O1} - \Delta\varepsilon_{Oid}$
and from $\Delta\varepsilon_{O2} - \Delta\varepsilon_{O1}$ as computed from the
values from both~\cite{Garcia2011a} and~\cite{Razo-Mejia2018}. Three of the four
values are within $\sim0.5~k_BT$ of the diagonal representing perfect agreement,
which is comparable to the $\sim\text{few}\times0.1~k_BT$ variability between
the independent determinations of the same quantities between~\cite{Garcia2011a}
and~\cite{Razo-Mejia2018}. The only outlier involves Oid measurements
from~\cite{Razo-Mejia2018}, and as the authors of~\cite{Razo-Mejia2018} note,
this is a difficult measurement of low fluorescence signal against high
background since Oid represses so strongly. We are therefore inclined to regard
the failure of this point to fall near the diagonal as a testament to the
difficulty of the measurement and not as a failure of our theory.
